Approximating the matrix exponential derivative of the CTMC generator matrix, with application to global SARS-CoV-2

### 2023-03-16 manuscript/main.tex
post viz figs visualized

We generate 10 million MCMC samples in this manner, saving 1 in 10,000 Markov chain states in order to conserve memory.  
\begin{itemize}
	\item ESS
	\item Figure 2
	\item Tree
\end{itemize}

In addition to these scientific questions of interest, we are interested in the performance of the first-order approximation as a surrogate gradient for HMC in such a high-dimensional setting. To isolate this performance, we fix the Bayesian bridge global-scale $\tau$ at 2.5$\times10^{-5}$.  We generate a Markov chain with 80,000 states, saving every tenth state and removing the first 1000 states as burn-in.  Despite the relatively small number of iterations, we observe large ESS that suggest satisfactory accuracy of the first-order approximation within the context of high-dimensional surrogate HMC.  The ESS for the three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ are 1,053, 1,343 and 498, respectively.  The median and minimum ESS for the 1,892 random-effects are 1,514 and 1,161, respectively.

>>>>

We generate 8 million MCMC samples in this manner, saving 1 in 10,000 Markov chain states, in order to guarantee a minimum effective sample size (ESS) greater than 100.  By far, the worst-mixing parameter  is the global scale $\tau$, which obtains an ESS of 185.  The three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ obtain ESS's of 721, 721 and 644, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 721 and 190, respectively.  We note that, after thinning and removing burn-in, the sample only consists of 721 states, so ESS of 721 implies negligible autocorrelation between samples.  Figure \ref{fig:post} presents

\begin{itemize}
	\item Figure 2
	\item Tree
\end{itemize}

In addition to these scientific questions of interest, we are interested in the performance of the first-order approximation as a surrogate gradient for HMC in such a high-dimensional setting. To isolate this performance, we fix the Bayesian bridge global-scale $\tau$ at 2.5$\times10^{-5}$.  We generate a Markov chain with 80,000 states, saving every tenth state and removing the first 1000 states as burn-in.  Despite the relatively small number of iterations, we observe large ESS that suggest satisfactory accuracy of the first-order approximation within the context of high-dimensional surrogate HMC.  The ESS for the three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ are 1,053, 1,343 and 498, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 1,514 and 1,161, respectively.
